---
title: "Smoking Prevalence Notebook for Sharing"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(janitor)
library(magrittr)
library(ggmap)
library(googlesheets)
library(ggrepel)
```
# (expect a prompt to authenticate with Google in console)
Note you will have to authenticate for the googlesheets package to work, but I have provided my google sheets keys listed below that should allow access to my google sheets. 

Here's some help with googlesheets if needed: https://cran.r-project.org/web/packages/googlesheets/vignettes/basic-usage.html

## Import data from Google Sheets

I copied and pasted from the orginal article to a google sheets because I didn't know a faster or easier way. 
The URL for the data and table of interest is: https://www.thelancet.com/action/showFullTableHTML?isHtml=true&tableId=tbl1&pii=S0140-6736%2817%2930819-X
The full article is: https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(17)30819-X/fulltext#
Here's the graph that inspired me: http://cdn.static-economist.com/sites/default/files/imagecache/1872-width/20170415_WOC921.png
Here's the tweet that started it all: https://twitter.com/datawookie/status/1142042679351681025

```{r}
data_from_lancet_key <- gs_key("1I8FKCUhb6WM-vNOL19inNAmWv0Guq_Q9dgdMVsquFOk")

world_pop_key <- gs_key("1CqdhlwlwXmKA-YEgBhZJECgPqUwv2_PmB2bWpBzyEo0")

country_codes_key <- gs_key("1sXQqMu4smGlP7p8vxL6fq7VZSU3Tz-kZB0GtiFFzAKI")

data_from_lancet <- gs_read_csv(data_from_lancet_key)

world_pop_data <- gs_read_csv(world_pop_key, skip = 4)

country_codes <- gs_read_csv(country_codes_key)
```



Cleaning of my data

```{r}
world_pop_data <- clean_names(world_pop_data)

data_clean <- clean_names(data_from_lancet)

#make sdi_level a factor from a character
data_clean$sdi_level %<>% factor

#separate point values from confidence intervals
data_2_tidy <- data_clean %>% separate(x2015_female_age_standardised_prevalence, into = c("female_prevalence_2015", "female_confidence_interval"), " ", extra = "merge") %>% separate(x2015_male_age_standardised_prevalence, into = c("male_prevalence_2015", "male_confidence_interval"), " ", extra = "merge") 
#replace middle dot with period
data_2_tidy$female_prevalence_2015 <- gsub("·", ".", data_2_tidy$female_prevalence_2015)

data_2_tidy$male_prevalence_2015 <- gsub("·", ".", data_2_tidy$male_prevalence_2015)

#repeating prior two steps for other columns
data_2_tidy_all_columns <- data_2_tidy %>% separate(annualised_rate_of_change_female_1990_2015, into = c("female_1990_2015", "female_1990_2015_confidence_interval"), " ", extra = "merge") %>% separate(annualised_rate_of_change_male_1990_2015, into = c("male_1990_2015", "male_1990_2015_confidence_interval"), " ", extra = "merge") %>% separate(annualised_rate_of_change_male_1990_2005, into = c("male_1990_2005", "male_1990_2005_confidence_interval"), " ", extra = "merge") %>% separate(annualised_rate_of_change_male_2005_2015, into = c("male_2005_2015", "male_2005_2015_confidence_interval"), " ", extra = "merge") %>% separate(annualised_rate_of_change_female_1990_2005, into = c("female_1990_2005", "female_1990_2005_confidence_interval"), " ", extra = "merge") %>% separate(annualised_rate_of_change_female_2005_2015, into = c("female_2005_2015", "female_2005_2015_confidence_interval"), " ", extra = "merge") 

#Removing middle dot from other columns
#figure out how to make this happen with lapply or something
data_2_tidy_all_columns$female_1990_2015 <- gsub("·", ".", data_2_tidy_all_columns$female_1990_2015)

data_2_tidy_all_columns$female_1990_2005 <- gsub("·", ".", data_2_tidy_all_columns$female_1990_2005)

data_2_tidy_all_columns$female_2005_2015 <- gsub("·", ".", data_2_tidy_all_columns$female_2005_2015)

data_2_tidy_all_columns$male_1990_2015 <- gsub("·", ".", data_2_tidy_all_columns$male_1990_2015)

data_2_tidy_all_columns$male_1990_2005 <- gsub("·", ".", data_2_tidy_all_columns$male_1990_2005)

data_2_tidy_all_columns$male_2005_2015 <- gsub("·", ".", data_2_tidy_all_columns$male_2005_2015)

#Huge issue and roadblock here. Got great help from R Ladies r-help channel
#dash encoding issue: copy paste from the data did not get intrepreted correctly, so I looked at the unicode for it and then used the code

data_2_tidy_all_columns$female_prevalence_2015 <- as.numeric(data_2_tidy_all_columns$female_prevalence_2015)
data_2_tidy_all_columns$male_prevalence_2015 <- as.numeric(data_2_tidy_all_columns$male_prevalence_2015)


data_2_tidy_all_columns$female_1990_2015 <- as.numeric(gsub('\U2212', "-", data_2_tidy_all_columns$female_1990_2015))

data_2_tidy_all_columns$female_1990_2005 <- as.numeric(gsub('\U2212', "-", data_2_tidy_all_columns$female_1990_2005))

data_2_tidy_all_columns$female_2005_2015 <- as.numeric(gsub('\U2212', "-", data_2_tidy_all_columns$female_2005_2015))

data_2_tidy_all_columns$male_1990_2015 <- as.numeric(gsub('\U2212', "-", data_2_tidy_all_columns$male_1990_2015))

data_2_tidy_all_columns$male_1990_2005 <- as.numeric(gsub('\U2212', "-", data_2_tidy_all_columns$male_1990_2005))

data_2_tidy_all_columns$male_2005_2015 <- as.numeric(gsub('\U2212', "-", data_2_tidy_all_columns$male_2005_2015))


#filter out confidence intervals (not really needed but easier to look at)
data_2_tidy_no_cis <- data_2_tidy_all_columns[, c(-4, -6, -8, -10, -12, -14, -16, -18)]



```


Adding some population data for possible later analysis

```{r}
merged_data <- merge(data_2_tidy_no_cis, world_pop_data, by.x = 1, by.y = 1)

merged_data_with_codes <- merge(merged_data, country_codes, by.x = 11, by.y = 3)
```
```

Analysis to make a version of the Economist Plot
```{r}
economist_analysis <- data_2_tidy_no_cis %>% mutate(male_estimate_1990 = male_prevalence_2015 / ((1+(male_1990_2015/100))^25)) %>% mutate(change_overall_male = male_prevalence_2015 - male_estimate_1990) %>% mutate(female_estimate_1990 = female_prevalence_2015 / ((1+(female_1990_2015/100))^25)) %>% mutate(change_overall_female = female_prevalence_2015 - female_estimate_1990)

economist_analysis <- economist_analysis %>% mutate(quadrant = case_when(change_overall_male > 0 & change_overall_female > 0   ~ "Both Increased",
                              change_overall_male <= 0 & change_overall_female > 0  ~ "M Decreased, W Increased",
                              change_overall_male <= 0 & change_overall_female <= 0 ~ "M Increased, W Decreased",
                              TRUE                                         ~ "Both Decreased"))
```

## Plotting

```{r}
plot_6 <- ggplot(economist_analysis, aes(x = change_overall_male, y = change_overall_female, color = quadrant, label = name)) + 
  geom_point() + 
  theme_minimal() + 
  geom_hline(yintercept=0) + 
  geom_vline(xintercept=0) +
  
    geom_text_repel(
    data          = subset(economist_analysis, change_overall_male > 6.5 | change_overall_male < -15 | change_overall_female >= 3.5 | change_overall_female < -10 ),
    
    segment.size  = 0.1,
    segment.color = "grey50") +
  
  labs(title = "Change in Men's and Women's Smoking Prevalence from 1990 to 2015", subtitle = "Inspired from Twitter and The Economist", caption = "Data from The Lancet") + labs(x = "Change in Men's Prevalence", y = "Change in Women's Prevalence") + theme(legend.position = "none") + annotate("text", x = 6.5, y = 4, label = "Both up")  + annotate("text", x = -6.5, y = -11, label = "Both down") + annotate("text", x = 5, y = -11, label = "Men up, Women down") +  annotate("text", x = -5, y = 6, label = "Men down, Women up")


plot_6
```
